WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-11 articles/2014/Article-WSClean/wsclean.tex
Work on article

\subsection{Discretization of w} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers in a simulated field. WSClean was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle, which would normally require $\sim195$ $w$-layers. Sources are 1~Jy (red circles), ghost sources are approximately 0.2~Jy. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce more ghost sources. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal (TODO this is in contrast to what Cornwell 2008 says -- mailed Cornwell to ask why). From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{lrrr}%
\hline\hline
& \textbf{WSClean} & \textbf{WSClean}    & \textbf{CASA} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (minimum w-layers/planes)}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & --- & 1.90 mJy\\
Computational time & 8.5 min & & 19.3 min\\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (128 w-layers/planes)}} \\
\hline
Source flux standard error & 1.39\% & & 2.08\% \\
RMS in residual image & 0.94 mJy & --- & 0.94 mJy \\
Computational time & 10.3 min & & 19.6 min \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 10\degree}} \\
\hline
Source flux standard error & 1.75\% & 2.92\% & 2.41 \%\\
RMS in residual image & 0.90 mJy & 1.05 mJy & 1.07 mJy \\
Computational time & 15.3 min & 6.6 min & 178.2 min \\
& & recentreTODO & \\
\hline\hline
\end{tabular}\end{center}\end{table}

@@

To asses the accuracy of WSClean and CASA, we have simulated a MWA observation without system noise, but with hundred sources of 1~Jy in a 20\degree diameter area. A unitary primary beam is assumed. We imaged the simulated set both with WSClean and CASA, using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. The Aegean program \citep{aegean-hancock-2012} is used to perform source detection on the produced images. Sidelobe noise of the residual 10~mJy source structures triggers a few false detections. These are ignored.

Table~\ref{tbl:accuracy-measurements} lists the measured RMS in the residual image and the standard errors of the source brightnesses as detected by Aegean. In all tests, WSClean performs slightly more accurate. At zenith, CASA produces a residual RMS twice as high as WSClean. This is caused by the fact that we keep the number of w-projection planes in w-projection equal to the number of w-layers in w-stacking, resulting in only 12 $w$-planes at zenith. This evidently has a stronger effect on the w-projection algorithm. However, the computational performance of the w-projection algorithm is hardly affected by the number of w-projection planes, and in practical situations one would always use more w-projection planes\footnote{In the zenith case CASA actually issues a warning to use more $w$-projection planes.}. When 128 w-projection planes are used in CASA, its residual RMS is equal to WSClean with 12 $w$-layers. We do not know why this parameter needs to be higher in CASA to reach the same RMS. In theory it should be enough, since Eq.~\eqref{eq:nwlayers-bound} results in $N_\textrm{wlayers}\gg 6$ for the top-left image pixel that is 26\degree away from the centre, while the sources are all within 10\degree. Also unexpected is that the source flux density becomes worse by increasing the number of planes. For WSClean both values stay approximately the same when the number of $w$-layers is increased.

In the $\textrm{ZA}=10\degree$ case, CASA is slightly less accurate. We have noticed this consistently at higher zenith angles, and it can actually be observed in the produced images, as shown in Fig.~\ref{fig:residuals}. TODO the following might be caused by a bug; check: We have also made an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is lower compared to the normal projection. This is due to the fact that the sources are further away from the phase centre, and sources have therefore on average higher $(l,m)$-values, thereby increasing the average $w$-correction. Nevertheless, the reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

@@

Many parameters affect the accuracy more than the observed differences between the methods. These include the number of major iterations, $w$-planes/layers, restoring beam, image size, weighting, kernel size and grid oversampling. Therefore, the differences in accuracy are not very significant. The results do show that synthesis imaging using common imaging parameters results in an approximate flux-density standard error of at least $\sim1\%$.

\begin{figure*}%
\begin{subfigure}{.36\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-nsamples/nsamples}%
\caption{}\label{fig:timing-nsamples}%
\end{subfigure}%
\begin{subfigure}{.36\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-zenith-angle/za}%
\caption{}\label{fig:timing-zenith-angle}%
\end{subfigure}\\
\begin{subfigure}{.36\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-resolution/resolution}
\caption{}\label{fig:timing-resolution}%
\end{subfigure}%
\hspace{-.05\linewidth}\begin{subfigure}{.36\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-channels/channels}
\caption{}\label{fig:timing-channels}%
\end{subfigure}%
\hspace{-.05\linewidth}\begin{subfigure}{.36\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-fov/fov}
\caption{}
\label{fig:timing-fov}
\end{subfigure}%
\caption{Imaging performance as a function of several configuration parameters. Error bars show 5$\sigma$ level. }\label{fig:timings}
\end{figure*}

@@

In Panel~\ref{fig:timing-zenith-angle}, the computational cost as function of zenith angle is plotted. It shows that, as expected from \S\ref{sec:time-complexity}, the $w$-projection algorithm is more affected by the increased $w$-values, leading to differences of more than an order of magnitude at zenith angles of $\gtrapprox 20$\degree. Additionally, at higher zenith angles, the $w$-kernels become too large to be able to make images of 3072 pixels or larger. Using the recentering technique to decrease the $w$-values makes the computational cost approximately constant. However, the additional time required to rephase and regrid the measurement set makes it only worthwhile for images $\ge3072$ pixels and $\textrm{ZA}\ge15\degree$. TODO: this is quite heavily depending on the speed of \texttt{chgcentre}, and that might still be improvable...

@@

Panel~\ref{fig:timing-channels} shows performance versus the spectral output resolution. The cost of imaging multiple frequencies is the cost of making an image from fewer visibilities times the number of frequencies. FFTs become the dominant cost for such small visibility sets, and the number of imaged frequencies affect performance linearly.

@@

\subsection{Derivation of computing cost formulae}
Based on the expected cost terms described in Sec.~\ref{sec:time-complexity}, we derive analytical functions for the CASA and WSClean measurements using least-squares fitting. Several functions with different free parameters were tested, and formulae with minimum amount of parameters are selected that still follow the trend of the measurements and have reasonably small errors. Measurements are weighted with the inverse standard deviation. The single outlying measurement of CASA in Fig.~\ref{fig:timing-nsamples} is removed.

For the WSClean time cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_e,N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_e + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_e,N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_e + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_e$ is the maximum $w$ value and is estimated with
\begin{equation}
 w_e = \frac{1}{\lambda}\left[\max |\Delta \mathbf{x}| \sin \textrm{ZA} + \max \Delta z \cos \textrm{ZA} \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right],
\end{equation}
where $\max |\Delta \mathbf{x}|$ is the maximum baseline length, $\max \Delta z$ is the maximum height difference between antennas and $\lambda$ is the wavelength.

TODO say something about error of analytical function vs measurements.

\begin{table*}
\caption{Predicted computational cost for various configurations, based on multi-frequency synthesis with 5 major iterations observing for 1 hour at ZA=20\degree.  Predictions are for $w$-projection in CASA, $w$-stacking with WSClean and $w$-stacked 12$\times$5-min snapshots with WSClean, with approximately equal accuracy.}
\begin{tabular}{lrrrrrrrrrr|rrr}
 Configuration&$\lambda$&FWHM  & Beams & Ant & Res & $\max |\Delta \mathbf{x}|$ & $\max \Delta z$ & int. t & BW & $\Delta \nu$ & $w$-proj & $w$-stack & $w$-snaps \\
          & (m)       &       &       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) &  &  & \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40     &14h 02m & 1h 52m & 1h 23m \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 &  0 &10 & 300 & 20    &19h 49m &14h 12m & 14h 13m\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 &    &10 & 16 & 16     & 2h 30m & 0h 11m & 0h 11m\\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 &    &10 & 16 & 16     & 2h 31m & 0h 38m & 0h 39m\\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 1 & 1 &  7 & 24     & 10h 47m & 4h 58m & 4h 58m \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0 &10 & 1.56 & 12.2 & 0h 32m & 0h 12m & 0h 02m\\
 SKA-P1 AA inner&3& 1.2\degree& 480  &  35 & 140'' &2.5& 2 &1.2 & 380 & 1     & 3218 d & 1865 d  & 1865 d \\
 SKA-P1 AA mid&3 & 1.2\degree & 480  &  50 & 3.5'' &100& 50 &1.2 & 380 & 1     & 6658 d & 3843 d & 3844 d \\
\end{tabular}
\end{table*}

@@

@@

\section{Application of WSClean to other telescopes} \label{sec:application-to-nonmwa}
TODO

@@

>>>>

\subsection{Discretization of w} \label{sec:gridding-w}
\begin{figure}
\begin{center}
\includegraphics[width=8cm]{img/aliasing-example-annotated}
\caption{Aliasing artefacts caused by insufficient $w$-layers in a simulated field. WSClean was set to use 12 $w$-layers. The centre of the image is at 10\degree zenith angle, which would normally require $\sim195$ $w$-layers. Sources are 1~Jy (red circles), ghost sources are approximately 0.2~Jy. Each source produces two ghost sources, but because they re-appear after a major cleaning cycle, they are eventually cleaned and produce more ghost sources. }
\label{fig:aliasing-example}
\end{center}
\end{figure}
While the discretization of $u$ and $v$ is similar to conventional imaging, the discretization of $w$ defines the number of w-layers that need to be processed. For this, one can use a rule similar to \eqref{eq:when-2d-is-valid}, and make sure that the phase difference for two subsequent discretized $w$-values, $w_A$ and $w_B$,  is less than one radian. This results in the constraint
\begin{equation} \label{eq:minimum-w-distance}
\left|\left(w_A - w_B\right) 2\pi (\sqrt{1-l^2-m^2}-1)\right| \ll 1.
\end{equation}
This suggest that a uniform discretization in $w$ is optimal (TODO this is in contrast to what Cornwell 2008 says -- Cornwell mailed me he made a mistake... Do we need to explicitly point that out here :-/ ?). From Eq.~\eqref{eq:minimum-w-distance}, the required number of layers can be derived and is given by
\begin{equation} \label{eq:nwlayers-bound}
 N_\textrm{wlayers} \gg 2\pi \left(w_e - w_0\right) \max_{l,m} \left(1 - \sqrt{1-l^2-m^2}\right).
\end{equation}
Actual values for the right-hand side can differ a lot. The value of $w_e - w_0$ is influenced by the coplanarity of the array, the zenith angle and the wavelength, while the value of the $\max_{l,m}$ term is influenced by the angular size of the image. For the MWA, a typical value of $w_e - w_0$ is $\sim 10$ at zenith, but is already $\sim 400$ at a zenith angle of $30^{\circ}$. For a typical full-field-of-view image of a MWA observation of $3072\times 3072$ pixels of $0.75''$ size, the $\max_{l,m}$ term is $0.68$. This implies that tens of $w$-layers are required at zenith and hundreds at lower elevations. The number of $w$-layers has a large effect on the performance of the $w$-stacking algorithm, and will be discussed further in the next sections.

@@

\subsection{Accuracy}
\begin{table}%
\caption{Results on imaging accuracy measurements.} \label{tbl:accuracy-measurements}%
\begin{center}\begin{tabular}{lrrr}%
\hline\hline
& \textbf{WSClean} & \textbf{WSClean}    & \textbf{CASA} \\
&                  & \textbf{+ recentre} & \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (6 w-layers/planes)}} \\
\hline
Source flux standard error & 1.31\% & & 1.34\% \\
RMS in residual image & 0.94 mJy & --- & 1.90 mJy\\
Computational time & 8.5 min & & 19.3 min\\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 0\degree (128 w-layers/planes)}} \\
\hline
Source flux standard error & 1.39\% & & 2.08\% \\
RMS in residual image & 0.94 mJy & --- & 0.94 mJy \\
Computational time & 10.3 min & & 19.6 min \\
\hline
\multicolumn{3}{l}{\textit{Zenith angle 10\degree}} \\
\hline
Source flux standard error & 1.75\% & 1.40\% & 2.41 \%\\
RMS in residual image & 0.90 mJy & 1.03 mJy & 1.07 mJy \\
Computational time & 15.3 min & 6.6 min & 178.2 min \\
\hline\hline
\end{tabular}\end{center}\end{table}

@@

To assess the accuracy of WSClean and CASA, we simulate a MWA observation with hundred sources of 1~Jy in a 20\degree diameter area, without adding system noise. A unitary primary beam is assumed. We image the simulated set with WSClean and CASA using Cotton-Schwab cleaning to a threshold of 10~mJy. The two imagers calculate slightly different restoring (synthesized) beams, hence to avoid bias the restoring beams are fixed. Other imaging parameters are given in Table~\ref{tbl:default-parameters}. The Aegean program \citep{aegean-hancock-2012} is used to perform source detection on the produced images. Sidelobe noise of the residual 10~mJy source structures triggers a few false detections. These are ignored.

Table~\ref{tbl:accuracy-measurements} lists the measured RMS in the residual image and the standard errors of the source brightnesses as detected by Aegean. In all tests, WSClean performs slightly more accurate. At zenith, CASA produces a residual RMS twice as high as WSClean. This is caused by the fact that we keep the number of $w$-projection planes in w-projection equal to the number of w-layers in w-stacking, resulting in only 12 $w$-planes at zenith. This evidently has a stronger effect on the $w$-projection algorithm. However, the computational performance of the $w$-projection algorithm is hardly affected by the number of $w$-projection planes, and in practical situations one would always use more w-projection planes\footnote{In the zenith case CASA issues a warning to use more $w$-projection planes.}. When 128 $w$-projection planes are used in CASA, its residual RMS is equal to WSClean with 12 $w$-layers, but the flux density measurements are less accurate. We do not know why this parameter needs to be higher in CASA to reach the same RMS.\footnote{TODO: do I want to find out this is because of invalid sqrt(w) gridding?} Eq.~\eqref{eq:nwlayers-bound} results in $N_\textrm{wlayers}\gg 6$ for the top-left image pixel that is 26\degree away from the centre, while the sources are all within 10\degree. Also unexpected is that the source flux density becomes worse by increasing the number of planes. For WSClean both values stay approximately the same when the number of $w$-layers is increased.

In the $\textrm{ZA}=10\degree$ case, CASA is slightly less accurate. The extra noise can be observed in the produced images, as shown in Fig.~\ref{fig:residuals}. Also made is an image with the technique of recentring a zenith phase-centred visibility set, as described in Sect.~\ref{sec:snapshot-imaging}. Its accuracy is higher compared to the normal projection, but its residual noise is also higher. The reprojection technique needs on average fewer $w$-layers to reach the same level of accuracy.

@@

\begin{figure*}%
\begin{subfigure}{.5\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-nsamples/nsamples}%
\caption{}\label{fig:timing-nsamples}%
\end{subfigure}%
\hspace{-.05\linewidth}\begin{subfigure}{.5\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-zenith-angle/za}%
\caption{}\label{fig:timing-zenith-angle}%
\end{subfigure}\\
\begin{subfigure}{.5\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-resolution/resolution}
\caption{}\label{fig:timing-resolution}%
\end{subfigure}%
\hspace{-.05\linewidth}\begin{subfigure}{.5\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-channels/channels}
\caption{}\label{fig:timing-channels}%
\end{subfigure}\\
\begin{subfigure}{.5\linewidth}%
\includegraphics[width=\linewidth]{img/benchmark-fov/fov}
\caption{}\label{fig:timing-fov}%
\end{subfigure}%
\caption{Imaging performance as a function of several parameters. Error bars show 5$\sigma$ level. Unfinished lines indicate the imager could not run succesfully.}\label{fig:timings}
\end{figure*}

@@

In Panel~\ref{fig:timing-zenith-angle}, the computational cost as function of zenith angle is plotted. It shows that, as expected from \S\ref{sec:time-complexity}, the $w$-projection algorithm is more affected by the increased $w$-values, leading to differences of more than an order of magnitude at zenith angles of $\gtrapprox 20$\degree. Additionally, at higher zenith angles, the $w$-kernels become too large to be able to make images of 3072 pixels or larger. Using the recentering technique to decrease the $w$-values makes the computational cost approximately constant. However, the additional time required to rephase and regrid the measurement set makes it only worthwhile for images $\ge3072$ pixels and $\textrm{ZA}\ge15\degree$. This is dependent on the speed of rotating the phase centre and regridding. Our program to change the phase centre is currently not multi-threaded, so its performance might be improvable.

@@

The cost of spectral imaging, i.e. imaging multiple frequencies, is the cost of making an image from fewer visibilities times the number of desired frequencies. Panel~\ref{fig:timing-channels} shows performance versus the spectral output resolution. As can be expected from the previous results, FFTs become the dominant cost for such small visibility sets, and the number of imaged frequencies affect performance linearly.

As can be seen in Panel~\ref{fig:timing-fov}, the field of view (FOV) has a large affect on performance when imaging off-zenith. This plot was created by varying the size of the pixel in the output image, such that the number of pixels in the image did not change. The FOV is calculated as angle between the left and right most pixel in the image. CASA is not able to make off-zenith images larger than $\sim30$\degree, and its imaging cost increases much more rapidly compared to the cost of WSClean, which implies that gridding is the major cost in this scenario. The image recentering technique is clearly beneficial for FOVs larger than $\sim30$\degree, and can even make a difference of an order of magnitude at very large FOVs of $\sim90$\degree.

@@

\subsection{Derivation of computing cost formulae}
Based on the expected cost terms described in Sec.~\ref{sec:time-complexity}, we derive analytical functions for the CASA and WSClean measurements using least-squares fitting. Several functions with different free parameters were tested, and formulae with minimum amount of parameters are selected that still follow the trend of the measurements and have reasonably small errors. Measurements are weighted with the inverse standard deviation. The single outlying measurement for $N_\textrm{pix}=1792$ with CASA in Fig.~\ref{fig:timing-resolution} is removed.

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_e,N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_e + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_e,N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_e + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_e$ is the maximum $w$ value and is estimated with
\begin{equation}
 w_e = \frac{1}{\lambda}\left[\max |\Delta \mathbf{x}| \sin \textrm{ZA} + \max \Delta z \cos \textrm{ZA} \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right],
\end{equation}
where $\max |\Delta \mathbf{x}|$ is the maximum baseline length, $\max \Delta z$ is the maximum height difference between antennas and $\lambda$ is the wavelength. These functions follow the trend of the measurements well, with an absolute error of 14.7\% and 20.7\% for respectively the WSClean and CASA functions.

@@

\begin{table*}
\caption{Predicted computational cost for various configurations, based on multi-frequency synthesis with 5 major iterations observing for 1 hour at ZA=20\degree.  Predictions are for $w$-projection in CASA, $w$-stacking with WSClean and $w$-stacked snapshots with WSClean and optimal snapshot duration, all with approximately equal accuracy.}
\begin{tabular}{lrrrrrrrrrr|rrr}
 Configuration&$\lambda$& FOV  & Beams & Ant & Res & $\max |\Delta \mathbf{x}|$ & $\max \Delta z$ & int. t & BW & $\Delta \nu$ & $w$-proj & $w$-stack & $w$-snaps \\
          & (m)       & (FWHM)&       &     &     & (km)                       & (m)             & (s)    & (MHz) &  (kHz) &  &  & \\
\hline
 GLEAM    &  2   & 24.7\degree& 1    & 128 & 2'   & 2.9 & 5 & 2 & 32 & 40     &11h 20m & 1h 49m & 1h 23m \\
 EMU      & 0.2  & 1\degree   & 30   &  36 & 10''  & 6 &  0 &10 & 300 & 20    &23h 52m &11h 52m & 11h 49m\\
 MSSS low &  5   & 9.8\degree & 5    &  20 & 100'' & 5 &    &10 & 16 & 16     & 0h 53m & 0h 12m & 0h 11m\\
 MSSS high&  2   & 3.8\degree & 5    &  40 & 120'' & 5 &    &10 & 16 & 16     & 1h 41m & 0h 39m & 0h 39m\\
 AARTFAAC &  5   & 45\degree  & 1    & 288 &  20'  & 0.3& 1 & 1 &  7 & 24     & 12h 01m & 4h 58m & 4h 58m \\
 VLSS     & 4.1  & 14\degree  & 1    &  27 &  80'' &11.1& 0 &10 & 1.56 & 12.2 & 0h 10m & 0h 11m & 0h 01m\\
 SKA-P1 AA inner&3& 1.2\degree& 480  &  35 & 140'' &2.5& 2 &1.2 & 380 & 1     & 3219 d & 1865 d  & 1865 d \\
 SKA-P1 AA mid&3 & 1.2\degree & 480  &  50 & 3.5'' &100& 50 &1.2 & 380 & 1     &  6740 d & 3845 d & 3844 d \\
\end{tabular}
\end{table*}

\subsection{Estimated computing cost for other telescopes} \label{sec:application-to-nonmwa}
We use the derived functions to estimate the computational cost of the algorithms for configurations of several telescopes. We estimate the cost of imaging an hour of ZA=20\degree data for the following surveys or array configurations: The Galactic and Extragalactic MWA survey (GLEAM); the Evolutionary Map of the Universe (EMU) ASKAP survey \citep{emu-norris-2011}; the low and high bands of the LOFAR Multifrequency Snapshot Sky Survey (MSSS)\footnote{See \url{https://www.astron.nl/radio-observatory/lofar-msss/lofar-msss}}; the Amsterdam—ASTRON Radio Transients Facility and Analysis Centre (AARTFAAC) project\footnote{See \url{http://www.aartfaac.org}}; The VLA Low-Frequency Sky Survey (VLSS, \citealt{vlss-2007}); and the low-frequency phase~1 aperture arrays of the Square-Kilometre Array, with inner (2.5~km) and mid (2.5~km) configurations \citep{ska-phase1-2010}. In multi-beam configurations such as the SKA and EMU configurations, we assume each beam is imaged separately, i.e., the FOV of a single beam is used and the computing cost is multiplied with the number of beams.

@@

@@

TODO The results do show that synthesis imaging using common imaging parameters results in an approximate flux-density standard error of at least $\sim1\%$.
